---
title: "individual1"
author: "Connor Hergenreter"
date: "April 17, 2019"
output: html_document
---

```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
names(acc)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
names(per)
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(readxl)
```

#Question 1

The following chart shows the amount of accidents that have occurred on each day. It shows that Satuday has the highest accident count at 5873. The "safest" day would be Wednesday at 4314 accidents.

```{r}
ggplot(acc, aes(x=DAY_WEEK)) + geom_bar(stat="count", fill="steelblue") + geom_text(aes(label=..count..), stat="count",position=position_stack(0.5)) + scale_x_discrete(limits=c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))
```

#Question 2

The chart below shows that the count of accidents by hour. The hour with the most amount of accidents is 6p.m. at 2000 and the hour with the least amount of accidents is 4a.m. at 813.

```{r}
ggplot(acc, aes(x=HOUR)) + geom_bar(stat="count", fill="steelblue") + geom_text(aes(label=..count..), stat="Count", position=position_stack(0.5)) + coord_cartesian(xlim=c(0,24))
```

#Question 3

The following chart shows the amount of drunk drivers per accident. The total count of accidents involving drunk drivers is 8769.

```{r}
ggplot(acc, aes(x=DRUNK_DR)) + geom_bar(stat="count", fill="steelblue") + geom_text(aes(label=..count..), stat="Count", position=position_stack(0.5))
```

#Part 3

#Question 1

Since we were also given county information I decided to do this on a county level. The resulting map shows that not all counties have had their accident information recorded. It appears that the most amount of crashes happen throughout southern United States with a good deal of them happening in costal areas.

```{r}
glcs <- read_xlsx("FRPP GLC United States.xlsx", skip=1)
counties <- map_data("county")

glcs$`State Code` <- as.numeric(glcs$`State Code`)
glcs$`County Code` <- as.numeric(glcs$`County Code`)
glcs$`City Code` <- as.numeric(glcs$`City Code`)
accGlcs <- left_join(acc, glcs, by = c("STATE" = "State Code", "COUNTY" = "County Code", "CITY" = "City Code"))

counties$region <- toupper(counties$region)
counties$subregion <- toupper(counties$subregion)
accident.map <- accGlcs %>% left_join(counties, by = c("State Name" = "region", "County Name" = "subregion"))
```

#Question 2

Using a facet wrap with month we are able to see accident history throughout the year. Turns out more crashes happen in summer, specifically in June and July, than the other seasons. I had originally thought winter would be the worst season, but even spring seems to have more crashes.

```{r}
ggplot(accident.map, aes(x=long, y=lat, fill=ST_CASE)) + geom_polygon(aes(group=group)) + labs(title="Accident Frequency in the United States by County", x="Longitude", y="Latitude")
                                                                                               
ggplot(accident.map, aes(x=long, y=lat, fill=ST_CASE)) + geom_polygon(aes(group=group)) + labs(title="Accident Frequency in the United States by County and by Month", x="Longitude", y="Latitude") + facet_wrap(~MONTH)
```