---
title: "individual1"
author: "Connor Hergenreter"
date: "April 17, 2019"
output: html_document
---

```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
names(acc)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
names(per)
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
```

#Question 1

The following chart shows the amount of accidents that have occurred on each day. It shows that Satuday has the highest accident count at 5873. The "safest" day would be Wednesday at 4314 accidents.

```{r}
ggplot(acc, aes(x=DAY_WEEK)) + geom_bar(stat="count", fill="steelblue") + geom_text(aes(label=..count..), stat="count",position=position_stack(0.5)) + scale_x_discrete(limits=c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))
```

#Question 2

The chart below shows that the count of accidents by hour. The hour with the most amount of accidents is 6p.m. at 2000 and the hour with the least amount of accidents is 4a.m. at 813.

```{r}
ggplot(acc, aes(x=HOUR)) + geom_bar(stat="count", fill="steelblue") + geom_text(aes(label=..count..), stat="Count", position=position_stack(0.5)) + coord_cartesian(xlim=c(0,24))
```

#Question 3

The following chart shows the amount of drunk drivers per accident. The total count of accidents involving drunk drivers is 8769.

```{r}
ggplot(acc, aes(x=DRUNK_DR)) + geom_bar(stat="count", fill="steelblue") + geom_text(aes(label=..count..), stat="Count", position=position_stack(0.5))
```